### 2015-11-12 ch02-qlogms/sysid.tex
Include bib items from Larsson2012
Gebrol in intro/conclusie qLogMS
Beunen in excitation design

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not know at the time of the design. There are two main solutions to this discussed in the engineering literature:
\begin{enumerate}
\item Iterative input design where the design updated as more information about the system is available, see e.g. \cite{Hjalmarsson2005,Gevers2005}.
\item Robust input design one tries to make the input design robust to uncertainty in the a priori knowledge of the true parameter values, see e.g. \cite{Rojas2007a} and references therein. 
\end{enumerate}

@@

%\subsection{Using prior system information}
%Consider a system $G(s,\theta)$ where some initial guess of the parameter vector,
% say $\hat{\theta}_0 \in \mathbb{R}^n$ exists. A D-optimal design can then be
% found based $\hat{\theta}_0$. The resulting determinant is
%\begin{align}
%  \left[\det\IF^\star\right]^{-1}
%    &= \min_{\phi_u(\omega)}
%             \left[\det\IF\right]^{-1}
%\end{align}
%
%The optimal input can be realized as sum of $n$ sinusoids.
%Such an input is typically not robust to the initial guess since \TODO{\ldots}.
%As a solution to this, the optimal input is modified so that a fraction of the
%power in the optimal design is taken away and redistributed over all frequencies.
%This ensures that some excitation power is available also at other frequencies.
%
%If $\lambda_i$ is the fraction of the input power at frequency $\omega_i$,
%then the determinant of $\IF^\star$ can be expressed as \Cite{Goodwin}
%\begin{align}
%  \det\IF^\star &= \prod_{i = 1}^{n}
%                      \lambda_i^2
%                      \alpha(\omega_i).
%\end{align}
%The optimal input spectrum is
%\begin{align}
%  \phi_u^\star(\omega) &= \sum_{i=1}^{n}
%                            \lambda_i^2
%                            \delta(\omega,\omega_i)
%\end{align}
%If a fraction, say $\eta$, of the power is taken from each frequency and used
%at other frequencies, the input spectrum can be written as
%\begin{align}
%  \phi_u(\omega) &= (1-\eta)
%                      \sum_{i=1}^{n}
%                        \lambda_i^2 
%                        \delta(\omega,\omega_i)
%                    + \phi_u^{\robust}(\omega),\\
%                 &= (1-\eta) \phi_u^\star    (\omega)
%                    +        \phi_u^{\robust}(\omega).
%\end{align}
%The term $ \phi_u^{\robust}$ is the part that introduces robustness and
%must be such that
%\begin{align}
%  \int_0^{\pi} \phi_u^{\robust}(\omega) \dd{\omega} &= n \eta.
%\end{align}
%The spectrum $\phi_u^{\robust}(\omega)$ can for example be chosen as
%a discrete spectrum, e.g. from a multi-sine signal, or as
%a continuous spectrum. The resulting information matrix for the robust input is hence
%\begin{align}
%  \IF &= (1-\eta)\IF^\star + \IF^\robust
%\end{align}
%The determinant is therefore
%\begin{multline}
%\det\left[(1-\eta)\IF^\star + \IF^\robust\right]
%  \geq \det\left[(1-\eta)\IF^\star\right] + \det\left[\IF^\robust\right] \\
%  =  (1-\eta)^n\det\left[\IF^\star\right] + \det\left[\IF^\robust\right]
%  \text{.}
%\end{multline}
%The inequality holds since the information matrix is a positive, semi-definite
%matrix. Hence the largest possible loss in the determinant due to the robustness
% is given by the factor $(1-\eta)^n$.
%\TODO{This bound is often very low, maybe better one should be found!}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y_t &= G(s,\theta)=\frac{1}{\frac{s^2}{\omega_0^2} + 2\frac{\damping}{\omega_0}s + 1}u_t,
\label{eq:2ndOrderSys}
\end{align}
with parameters $\theta=[\omega_0\;\damping]^{\T}$. The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\omega_0$
and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wR$.
Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep{Goodwin1977} we have that the D-optimal design for \eqref{eq:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac{\omega_0\left(\left(1-2\damping^2\right)+\left(\left(1-2\damping^2\right)^2+15\right)^{1/2}\right)^{1/2}}{\sqrt{5}}.
\label{eq:optFreq}
\end{align}
For systems with low damping, i.e $\damping\ll1$ , $\omega^\star\approx\omega_0$, which is inside the bandwidth of the system. From \eqref{eq:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. Input design can be based on a guess of $\omega_0$, however if this guess is wrong, the resulting estimates may be unacceptable \citep{Rojas2007}.

When no information about the system is available, we propose using an input power spectrum with a logarithmic power distribution. This can be achieved either with a continuous bandlimited $1/f$ spectrum or a quasi-logarithmic multisine spectrum \citep{Pintelon2001}. We choose to work with an equidistant frequency grid with resolution $f_0$ and design the input in the frequency band $[\omega_0k_{min}\; \omega_0k_{max}]$, where the resolution is $k_0=1/T_{meas}$ and $T_{meas}$ is the measurement time. Hence, a finite measurement time limits the frequency resolution and therefore we have to put a lower limit on the frequency band of interest. Moreover, we can only impose the logarithmic distribution sufficiently well if $k_{min}$ is large enough.

First, we look at the bandlimited $1/f$ power spectrum given by
\begin{align*}
\label{eq:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\omega_0k_{min}\; \omega_0k_{max}]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wR}^{(1+\damping)\wR}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\omega_0$. The robustness of the $1/f$ input has been noted previously in e.g. \cite{Rojas2007}

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2001}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\delta(\omega,\omega_k)},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha_{f} = e^{\Delta}$.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
\left\lfloor\ln{\left[\frac{1+\damping}{1-\damping}\right]}\Delta^{-1}\right\rfloor,
\end{align}
where $\lfloor\cdot\rfloor$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\omega_0$.

@@

The result of these two choices of input power spectrum is that for a given damping, $\var{\hat{G}}$ is approximately the same independent of $\omega_0$. 
This is illustrated in Section~\ref{sec:sim}.

\subsection{Sums of second order systems}
This section extends the previous section to more general systems than second order.
Consider a system
\begin{align}
G(s,\theta) &= \sum_{i=1}^{K} G_i(s,\theta_i)
             =\sum_{i=1}^{K}  \frac{1}
                                   {   \frac{s^2}{\omega_i^2}
                                    + 2\frac{\damping_i}{\omega_i}s
                                    + 1},\\
\theta_i&=[\omega_i\;\damping_i]^{\T}.
\label{eq:sumSys}
\end{align}
The systems  $G_i$ will be called the \emph{sub-systems} of $G$. With loss of generality, we assume that $\omega_1 < \omega_2 < \cdots < \omega_K$. To properly identify the system $G$, excitation is required such that all
subsystems are properly excited, regardless of their individual resonance frequencies.

@@

\subsubsection{Example: Sum of two second order systems}
Consider the system
\begin{align}
G(s,\theta) &= G_1(s,\theta_1) + G_2(s,\theta_2)%\\
%&=\frac{1}{\frac{s^2}{\omega_1^2} + 2\frac{\damping_1}{\omega_1}s + 1} + \frac{1}{\frac{s^2}{\omega_2^2} + 2\frac{\damping_2}{\omega_2}s + 1},
\end{align}
with $\theta_1=[\omega_1\;\damping_1]^{\T}$ and $\theta_2=[\omega_2\;\damping_2]^{\T}$.
The information matrix for this system is
\begin{align}
\IF &=
       \begin{bmatrix}
         \partial G_{1,1}  & \partial G_{1,2}\\
         \partial G_{2,1}  & \partial G_{2,2}
       \end{bmatrix}.
\label{eq:infoEx}
\end{align}
To investigate the decoupling, we study the $(1,1)$ elements of blocks $\partial G_{1,1}$ and $\partial G_{1,2}$ of the information matrix \eqref{eq:infoEx}, when the input spectrum \eqref{eq:pink} is used; similar arguments can be made for the other elements of the two matrix blocks.

Direct calculations for $\damping < 1$ give
\begin{align}
\label{eq:auto}
\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_1^{\HT}}{\partial \omega_1}\right\}\frac{1}{\omega}\dd\omega = \frac{\pi}{16\omega_1^2}\damping^{-3} + \mathcal{O}(\damping^{-1}).
\end{align}
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
% \begin{align}
% &\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right\}\frac{1}{\omega}\dd\omega\right| \leq\label{eq:cross1} \\
% &\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right\}\right|\frac{1}{\omega}\dd\omega \leq \\
% &\int_0^\infty\left|\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right|\frac{1}{\omega}\dd\omega.\label{eq:cross}
% \end{align}
\begin{gather}
\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:cross1} \\
\leq
\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \omega_1}\frac{\partial G_2^{\HT}}{\partial \omega_2}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \omega_1} &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\omega_1\damping}\right]}{\omega_1^2\left[1+2\frac{\damping j\omega}{\omega_1} + \left(\frac{j\omega}{\omega_1}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\omega_1\right\}$. Hence $\left|\frac{\partial G_1}{\partial \omega_1} \right|$ has asymptotical slope 20\unit{\frac{dB}{dec.}} for low frequencies, at $j\omega=\damping\omega_1$ the slope increases to 40\unit{\frac{dB}{dec.}}. There are four poles which give a $-40$\unit{\frac{dB}{dec.}} roll-off after the peak. The shape of $\frac{\partial G_1}{\partial \omega_1} $ is illustrated in Fig.~\ref{fig:dG}.

A crude, albeit useful, bound on the integrand in \eqref{eq:cross} is
\begin{align}
&f\left(\omega,\theta_1,\theta_2\right)=
  \begin{cases}
   \frac{\omega^2}{\omega_1^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \leq \omega_1$,}\\
   \phantom{\frac{\omega^2}{\omega^2}}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega_1 < \omega < \omega_2$,}\\
   \frac{\omega_2^2}{\omega^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \geq \omega_2$,}
  \end{cases}
\end{align}
where $\partial G_1^\infty = \left\|\frac{\partial G_1}{\partial \omega_1}\right\|_\infty$. Again, direct calculations give
\begin{align}
\int_0^\infty f\left(\omega,\theta_1,\theta_2\right)\dd\omega = \frac{\omega_1\left(2-\ln{\frac{\omega_1^2}{\omega_2^2}}\right)}{\omega_2^2}\damping^{-2} + \mathcal{O}(1).
\end{align}
Hence \eqref{eq:cross1} cannot grow faster than $\mathcal{O}(\damping^{-2})$ as $\damping\rightarrow0$. From this analysis we conclude that:
\begin{itemize}
\item the integral \eqref{eq:cross1} decreases as $\omega_2-\omega_1$ increases, and
\item lower damping further decreases the integral \eqref{eq:cross1} in comparison to \eqref{eq:auto}.
\end{itemize}

\begin{figure}
\centering
\input{\thisDir/figs/rolloff.tikz}
\caption{The shape of $\frac{\partial G_1}{\partial \omega_1}$. The roll-off around the peak is significant and depends on the damping of the system.}
\label{fig:dG}
\end{figure}

To further illustrate the decoupling of the two systems, we calculate \eqref{eq:infoEx} for different values of the damping $\damping$ and separations of the two resonances $\omega_1$ and $\omega_2$. If the separation is taken in terms of $\wdB$-units for the lower resonance, the result only depends on the difference $\omega_1-\omega_2$ and not on the actual frequencies. The resulting information matrix is then normalized so that the largest element is $1$.

@@

\begin{figure}[hpt]
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{5cm}
  \input{\thisDir/figs/Sys005log0-1.tikz}
  \caption{FRF of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.05$ 
           along with its variance and bias.}%
  \label{fig:FRF1}
\end{figure}

@@

\subsection{Simulation example on higher order systems}
To show that the suggested method also works on more general systems,
we considered a system that consists of two second order systems \eqref{eq:tf} in 
parallel. This is equivalent to adding their respective transfer functions,
therefore, the resulting transfer function is of the form:
\begin{equation}
  G \left( z \right) 
  = \frac{         b_4 z^{4} + b_3 z^{3} + b_2 z^{2} + b_1 z + b_0}%
         {\phantom{a_4}z^{4} + a_3 z^{3} + a_2 z^{2} + a_1 z + a_0}
  \text{.}
\label{eq:tfO4}
\end{equation}

For example, a system composed of two second order systems with $\damping = 0.2$
and $\fR = 400 \unit{Hz}$ for the first system and $\fR=4 \unit{kHz}$ for the second
system was simulated similarly to the previous examples. The standard deviation
of the noise $\sigma_e$ was put at $0.0678$. The corresponding results are 
shown in \figref{fig:sumSys}.

\begin{figure}%
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{5cm}
  \input{\thisDir/figs/Sum02log0-2.tikz}
  \caption{FRF for the sum of two second order systems with $\damping=0.2$ and
           resonances at $ 400 \unit{Hz}$ and 4 \unit{kHz}.}%
  \label{fig:sumSys}
\end{figure}

We observe that the amplitude of the FRF is $9.8\unit{dB}$ and $8.1\unit{dB}$ at the resonant frequencies.
The standard deviation on the FRF lies respectively $59.7 \unit{dB}$ and $59.6 \unit{dB}$ lower than the FRF at these frequencies. 
This shows that both resonances are captured with a practically identical error.

>>>>

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not know at the time of the design. There are two main solutions to this discussed in the engineering literature:
\begin{enumerate}
\item Iterative input design where the design updated as more information about the system is available, see e.g. \cite{Hjalmarsson2005,Gevers2005}.
\item Robust input design one tries to make the input design robust to uncertainty in the a priori knowledge of the true parameter values, see e.g. \cite{Rojas2007} and references therein. 
\end{enumerate}

@@

%\subsection{Using prior system information}
%Consider a system $G(s,\theta)$ where some initial guess of the parameter vector,
% say $\hat{\theta}_0 \in \mathbb{R}^n$ exists. A D-optimal design can then be
% found based $\hat{\theta}_0$. The resulting determinant is
%\begin{align}
%  \left[\det\IF^\star\right]^{-1}
%    &= \min_{\phi_u(\omega)}
%             \left[\det\IF\right]^{-1}
%\end{align}
%
%The optimal input can be realized as sum of $n$ sinusoids.
%Such an input is typically not robust to the initial guess since \TODO{\ldots}.
%As a solution to this, the optimal input is modified so that a fraction of the
%power in the optimal design is taken away and redistributed over all frequencies.
%This ensures that some excitation power is available also at other frequencies.
%
%If $\lambda_i$ is the fraction of the input power at frequency $\wn[i]$,
%then the determinant of $\IF^\star$ can be expressed as \Cite{Goodwin}
%\begin{align}
%  \det\IF^\star &= \prod_{i = 1}^{n}
%                      \lambda_i^2
%                      \alpha(\wn[i]).
%\end{align}
%The optimal input spectrum is
%\begin{align}
%  \phi_u^\star(\omega) &= \sum_{i=1}^{n}
%                            \lambda_i^2
%                            \delta(\omega,\wn[i])
%\end{align}
%If a fraction, say $\eta$, of the power is taken from each frequency and used
%at other frequencies, the input spectrum can be written as
%\begin{align}
%  \phi_u(\omega) &= (1-\eta)
%                      \sum_{i=1}^{n}
%                        \lambda_i^2 
%                        \delta(\omega,\wn[i])
%                    + \phi_u^{\robust}(\omega),\\
%                 &= (1-\eta) \phi_u^\star    (\omega)
%                    +        \phi_u^{\robust}(\omega).
%\end{align}
%The term $ \phi_u^{\robust}$ is the part that introduces robustness and
%must be such that
%\begin{align}
%  \int_0^{\pi} \phi_u^{\robust}(\omega) \dd{\omega} &= n \eta.
%\end{align}
%The spectrum $\phi_u^{\robust}(\omega)$ can for example be chosen as
%a discrete spectrum, e.g. from a multi-sine signal, or as
%a continuous spectrum. The resulting information matrix for the robust input is hence
%\begin{align}
%  \IF &= (1-\eta)\IF^\star + \IF^\robust
%\end{align}
%The determinant is therefore
%\begin{multline}
%\det\left[(1-\eta)\IF^\star + \IF^\robust\right]
%  \geq \det\left[(1-\eta)\IF^\star\right] + \det\left[\IF^\robust\right] \\
%  =  (1-\eta)^n\det\left[\IF^\star\right] + \det\left[\IF^\robust\right]
%  \text{.}
%\end{multline}
%The inequality holds since the information matrix is a positive, semi-definite
%matrix. Hence the largest possible loss in the determinant due to the robustness
% is given by the factor $(1-\eta)^n$.
%\TODO{This bound is often very low, maybe better one should be found!}

@@

 \subsection{A second order system}
Consider the second order system
\begin{align}
y_t &= G(s,\theta)=\frac{1}{\frac{s^2}{\wn^2} + 2\frac{\damping}{\wn}s + 1}u_t,
\label{eq:2ndOrderSys}
\end{align}
with parameters $\theta=[\wn\;\damping]^{\T}$. The resonance frequency of the system is $\wR \isdef \sqrt{1-\damping^2}\wn$
and the $3\unit{dB}$ bandwidth of the system is $\wdB \isdef 2\damping\wR$.
Since information is local in the frequency domain, the input signal power inside $\wdB$, contributes significantly to the information matrix while input power outside $\wdB$ has a smaller contribution. This can be seen by considering the optimal design for a second order system. From \citep{Goodwin1977} we have that the D-optimal design for \eqref{eq:2ndOrderSys} is a single sinusoid at frequency
\begin{align}
\omega^\star =\frac%
{\wn
  \sqrt{
    \left(1-2\damping^2\right)
    +
    \left(\left(1-2\damping^2\right)^2+15\right)^{1/2}
    }
  }
{\sqrt{5}}.
\label{eq:optFreq}
\end{align}
For systems with low damping, i.e $\damping\ll1$ , $\omega^\star\approx \wn$, which is inside the bandwidth of the system. 
From \eqref{eq:optFreq} it is clear that the optimal design depends on the resonance of the system to be identified. 
Input design can be based on a guess of $\wn$, however if this guess is wrong, the resulting estimates may be unacceptable~\citep{Rojas2007}.

When no information about the system is available, we propose using an input power spectrum with a logarithmic power distribution. This can be achieved either with a continuous bandlimited $1/f$ spectrum or a quasi-logarithmic multisine spectrum \citep{Pintelon2001}. We choose to work with an equidistant frequency grid with resolution $f_0$ and design the input in the frequency band $[\wn \kMin\; \wn \kMax]$, where the resolution is $k_0=1/T_{meas}$ and $T_{meas}$ is the measurement time. Hence, a finite measurement time limits the frequency resolution and therefore we have to put a lower limit on the frequency band of interest. Moreover, we can only impose the logarithmic distribution sufficiently well if $\kMin$ is large enough.

First, we look at the bandlimited $1/f$ power spectrum given by
\begin{align*}
\label{eq:pink}
\phi_u(\omega) &=\begin{cases}\frac{1}{|\omega|} &\text{if $|\omega|\in[\wn \kMin\; \wn \kMax]$,}\\
                        0           &\text{otherwise}.
            \end{cases}
\end{align*}
With this input spectrum, the power inside $\wdB$ is
\begin{align}
  \frac{1}{2\pi}
    \int\limits_{(1-\damping)\wn}^{(1+\damping)\wn}
     \frac{1}{\omega} \dd{\omega}
  = \frac{1}{2\pi}
    \ln{\frac{1+\damping}{1-\damping}},
\end{align}
which is independent of $\wn$. The robustness of the $1/f$ input has been noted previously in e.g. \cite{Rojas2007}

Second, we consider the power spectrum of the quasi-logarithmic multisine signal [\cite{Pintelon2001}] given by
\begin{align}
  \phi_u(\omega) &= \frac{1}{M}\sum_{k=1}^{M}{\delta(\omega,\wn[K])},
\end{align}
where $\omega_{k}/\omega_{k+1} \approx \alpha_{f} = e^{\Delta}$.
Hence, the number of spectral lines inside \wdB is 
\begin{align}
\left\lfloor\ln{\left[\frac{1+\damping}{1-\damping}\right]}\Delta^{-1}\right\rfloor,
\end{align}
where $\lfloor\cdot\rfloor$ is the nearest integer rounded towards $0$. The number of spectral lines is independent of $\wn$.

@@

The result of these two choices of input power spectrum is that for a given damping, $\var{\hat{G}}$ is approximately the same independent of $\wn$. 
This is illustrated in Section~\ref{sec:sim}.

\subsection{Sums of second order systems}
This section extends the previous section to more general systems than second order.
Consider a system
\begin{align}
G(s,\theta) &= \sum_{i=1}^{K} G_i(s,\theta_i)
             =\sum_{i=1}^{K}  \frac{1}
                                   {   \frac{s^2}{\wn[i]^2}
                                    + 2\frac{\damping_i}{\wn[i]}s
                                    + 1},\\
\theta_i&=[\wn[i]\;\damping_i]^{\T}.
\label{eq:sumSys}
\end{align}
The systems  $G_i$ will be called the \emph{sub-systems} of $G$. 
Without loss of generality, we assume that $\wn[1] < \wn[2] < \cdots < \wn[K]$. 
To properly identify the system $G$, excitation is required such that all subsystems are properly excited, regardless of their individual resonance frequencies.

@@

\subsubsection{Example: Sum of two second order systems}
Consider the system
\begin{align}
G(s,\theta) &= G_1(s,\theta_1) + G_2(s,\theta_2)%\\
%&=\frac{1}{\frac{s^2}{\wn[1]^2} + 2\frac{\damping_1}{\wn[1]}s + 1} + \frac{1}{\frac{s^2}{\wn[2]^2} + 2\frac{\damping_2}{\wn[2]}s + 1},
\end{align}
with $\theta_1=[\wn[1]\;\damping_1]^{\T}$ and $\theta_2=[\wn[2]\;\damping_2]^{\T}$.
The information matrix for this system is
\begin{align}
\IF &=
       \begin{bmatrix}
         \partial G_{1,1}  & \partial G_{1,2}\\
         \partial G_{2,1}  & \partial G_{2,2}
       \end{bmatrix}.
\label{eq:infoEx}
\end{align}
To investigate the decoupling, we study the $(1,1)$ elements of blocks $\partial G_{1,1}$ and $\partial G_{1,2}$ of the information matrix \eqref{eq:infoEx}, when the input spectrum \eqref{eq:pink} is used; similar arguments can be made for the other elements of the two matrix blocks.

Direct calculations for $\damping < 1$ give
\begin{align}
\label{eq:auto}
\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_1^{\HT}}{\partial \wn[1]}\right\}\frac{1}{\omega}\dd\omega = \frac{\pi}{16\wn[1]^2}\damping^{-3} + \mathcal{O}(\damping^{-1}).
\end{align}
For the corresponding element in $\partial G_{1,2}$ we start by bounding the absolute value of the integral by
% \begin{align}
% &\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right| \leq\label{eq:cross1} \\
% &\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega \leq \\
% &\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega.\label{eq:cross}
% \end{align}
\begin{gather}
\left|\int_0^\infty\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\frac{1}{\omega}\dd\omega\right|
\qquad \qquad \qquad \qquad \qquad
\label{eq:cross1} \\
\leq
\int_0^\infty\left|\mathrm{Re}\left\{\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right\}\right|\frac{1}{\omega}\dd\omega
\leq \\
\qquad \qquad \qquad \qquad \qquad
\int_0^\infty\left|\frac{\partial G_1}{\partial \wn[1]}\frac{\partial G_2^{\HT}}{\partial \wn[2]}\right|\frac{1}{\omega}\dd\omega \text{.}\label{eq:cross}
\end{gather}
To proceed, consider
\begin{align}
\frac{\partial G_1}{\partial \wn[1]} &= \frac{2\damping j\omega\left[1+\frac{j\omega}{\wn[1]\damping}\right]}{\wn[1]^2\left[1+2\frac{\damping j\omega}{\wn[1]} + \left(\frac{j\omega}{\wn[1]}\right)^2\right]^2},
\end{align}
which has zeros in $j\omega=\left\{0, \damping\wn[1]\right\}$. 
Hence $\left|\frac{\partial G_1}{\partial \wn[1]} \right|$ has asymptotic slope 20\unit{dB/\decade} for low frequencies, at $j\omega=\damping\wn[1]$ the slope increases to 40\unit{dB/\decade}. There are four poles which give a $-40$\unit{dB/\decade} roll-off after the peak. 
The shape of $\frac{\partial G_1}{\partial \wn[1]} $ is illustrated in Fig.~\ref{fig:dG}.

A crude, albeit useful, bound on the integrand in \eqref{eq:cross} is
\begin{align}
&f\left(\omega,\theta_1,\theta_2\right)=
  \begin{cases}
   \frac{\omega^2}{\wn[1]^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \leq \wn[1]$,}\\
   \phantom{\frac{\omega^2}{\omega^2}}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\wn[1] < \omega < \wn[2]$,}\\
   \frac{\wn[2]^2}{\omega^2}
   \frac{1}{\omega}
   \partial G_1^\infty
                                       & \text{if $\omega \geq \wn[2]$,}
  \end{cases}
\end{align}
where $\partial G_1^\infty = \left\|\frac{\partial G_1}{\partial \wn[1]}\right\|_\infty$. 
Again, direct calculations give
\begin{align}
\int_0^\infty f\left(\omega,\theta_1,\theta_2\right)\dd\omega = \frac{\wn[1]\left(2-\ln{\frac{\wn[1]^2}{\wn[2]^2}}\right)}{\wn[2]^2}\damping^{-2} + \mathcal{O}(1).
\end{align}
Hence \eqref{eq:cross1} cannot grow faster than $\mathcal{O}(\damping^{-2})$ as $\damping\rightarrow0$. From this analysis we conclude that:
\begin{itemize}
\item the integral \eqref{eq:cross1} decreases as $\wn[2]-\wn[1]$ increases, and
\item lower damping further decreases the integral \eqref{eq:cross1} in comparison to \eqref{eq:auto}.
\end{itemize}

\begin{figure}
\centering
\input{\thisDir/figs/rolloff.tikz}
\caption{The shape of $\frac{\partial G_1}{\partial \wn[1]}$. The roll-off around the peak is significant and depends on the damping of the system.}
\label{fig:dG}
\end{figure}

To further illustrate the decoupling of the two systems, we calculate \eqref{eq:infoEx} for different values of the damping $\damping$ and separations of the two resonances $\wn[1]$ and $\wn[2]$. If the separation is taken in terms of $\wdB$-units for the lower resonance, the result only depends on the difference $\wn[1]-\wn[2]$ and not on the actual frequencies. The resulting information matrix is then normalized so that the largest element is $1$.

@@

\begin{figure}[hpt]
  \centering
  \setlength{\figurewidth}{0.8\columnwidth}
  \setlength{\figureheight}{5cm}
  \input{\thisDir/figs/Sys005log0-1.tikz}
  \caption{FRF of a second order system with $\fR=40\unit{Hz}$ and $\damping=0.05$ 
           along with its variance and bias.}%
           \TODO{reproduce figure with ALL excitation signals}
  \label{fig:FRF1}
\end{figure}

@@
